Cultural Evolutionary Dynamics with Success and Influence biases

### 2022-06-12 ms/PrestigeBias.tex
update

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/chi_square_stats/GBD_validation.pdf}
  \caption{
  \textbf{Numerical validation for the GBD approximation.}
  The approximation (orange) fits simulation results (blue) well when using 1,000 simulations for both models.
  Here, population size, $N=2,000$; bias weight, $\alpha=0.1$; idea phenotype value, $\hat{A}=1$; role-model traits $\vec{A} \sim N(0,1)$; success bias value, $\beta(A)=0.956$.}	
  \label{fig:GBD}
\end{figure}

@@

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/chi_square_stats/DM_validation.pdf}
  \caption{
  \textbf{Goodness of fit of Dirichlet-Multinomial approximation to the model.}
  The difference between the DM approximation distribution to observed model distribution (a) is very small, with the highest difference at 0.5 role-models.
  The optimal negative log likelihood (b) is given when using the DM approximation (black vertical line), and is much better when compared to a shuffled vector of the DM distribution (blue bars).
  Here, population size, $N=100$; number of distributions averaged, $m=100$; phenotype values, $\hat{A}=1$, $A \sim N(0,1)$; homogeneous success bias weight, $\alpha=0.5$.}	
  \label{fig:DM_validation}
\end{figure}

@@

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/binary/num_sims.pdf}
  \caption{
  \textbf{Minimum number of simulations required to get a good approximation.}
  The approximation (orange) fits simulation results (blue) well when using 1,000 simulations for both models. The approximated value (black) is given by equation \cref{eq:kimura_p}.
  Markers for average value across simulations, error bars for 95\% confidence interval.
  Here, population size, $N=1000$; bias weight, $\alpha=0.5$; phenotype values, $\hat{A}=1$, $A=0.7$; success bias value, $\beta(A)=0.956$.}	
  \label{fig:num_sims}
\end{figure}

\begin{figure}
    \includegraphics[width=\linewidth]{../figures/binary/full_vs_dm_mutation.pdf}
  \caption{
  \textbf{Robustness of DM approximations to inclusion of estimation error.} Both the DM approximation (orange) and Kimura's equation (black line) fit the stochastic simulations (blue) well even with a high estimation error rate. Markers for average across simulations, error bars for 95\% confidence intervals.
  5,000 simulations per data point; population size, $N=1000$; bias weight, $\alpha=0.1$; phenotype values, $\hat{A}=1$,$A=0.7$; bias strength parameter $J\sim N(1,x^2)$ where $x \in [10^{-7},10^{-1}]$.
  }	
  \label{fig:hetro_error}
\end{figure}

\begin{figure}
    \includegraphics[width=\linewidth]{../figures/binary/full_vs_dm_changing_alpha.pdf}
   \caption{\textbf{Robustness of DM approximations to variation in the bias weight $\alpha$.} 
   Both the DM approximation (orange) and Kimura's equation (black line) fit the stochastic simulations (blue) well even with a high variation in success bias weight.Markers for average across $5,000$ simulations, error bars are 95\% confidence intervals.
  Here, population size, $N=1000$; success bias weight normally distributed, $\alpha\sim N(0.5,x^2)$; phenotype values ,$\hat{A}=1$,$A=0.7$; success bias value, $\beta(A)=0.956$.}	
  \label{fig:hetro_alpha}
\end{figure}

@@

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/binary/kimura_var.pdf}
    % TODO - black line for 2s as well
  \caption{\textbf{Fixation probability and time approximation for varying bias weight and success coefficient.}
  Fixation probability (\textbf{A}) and time (\textbf{B}) for varying bias weight on the bottom x-axis. The corresponds effective population size is on the top x-axis.
  The approximation (black) fits DM simulations results (green) and Wright-Fisher simulation results (orange) when using \cref{eq:kimura_p}.
  Fixation probability (\textbf{A}) is limited at approximately $2s$ (blue), as in the classic WF model.
  Markers are averages of $10,000$ simulations, error bars show 95\% confidence intervals for (\textbf{A,B}) and 75\% for (\textbf{C,D}).
   Here, Population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=0.7$ (\textbf{A,B}), $A=a\cdot \hat{A}, a \in [0.01,0.99]$ (\textbf{C,D}); success coefficient, $1-\beta=s=0.044$ (\textbf{A,B}); bias weight, $\alpha=0.01$ (\textbf{C,D}).}
  \label{fig:var_alpha}
\end{figure}

@@

\begin{figure}[H]
    \includegraphics[width=\linewidth]{../figures/changed_env/ch_env.pdf}
  \caption{\textbf{Sensitivity of fixation probability approximation in a changing environment to varying success coefficient and bias weight and robustness to varying ratios of cycles.}
\textbf{(A)} When the bias weight, or effective population size varies, the approximation (blue; \cref{eq:ch_env}) fits simulation results (orange). 
\textbf{(B)} When success bias is large ($1-\beta > 0.1$),  
simulation results (orange) are underestimated by the approximation (blue; \cref{eq:ch_env}). With even larger success bias ($1-\beta > 0.35$), even the constant environment approximation (green; \cref{eq:kimura_p}) underestimate simulation results, likely because the diffusion equation approximation assumes weak selection.
We can also see the robustness of the approximation to changes in cycle ratios (\textbf{C,D}): The approximation (blue) fits the simulations (orange) well for different sizes of the changing environment cycle, where $k$ is the number of generations where the bias is advantageous, and $l$ are the generations it is disadvantageous.
When $k<l$ the approximation is good. When $k>l$, the approximation and the simulations are both very close to the constant environment approximation (green). 
Markers show average of $10,000$ simulations, error bars show are 75\% \textbf{(A,C,D)} and 95\% \textbf{(B)} confidence intervals.
  Here, population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=0.9$ (\textbf{A,B}), $A=0.8$ \textbf{(C,D)}; In \textbf{(A)} success bias/selection coefficient is: $1-\beta=s=0.005$; In \textbf{(B,C,D)} success bias weight is: $\alpha=0.1$.}
  \label{fig:ch_env_alpha_beta}
\end{figure}

@@

\begin{figure}[H]
    \includegraphics[width=\linewidth]{../figures/influence_advantage/choose_bias.pdf}
    % TODO - capital letters for all axis lables
    % letter follows the description (success-bias weight, \alpha
    % show plot of error bars and a scatter for the constant bias weight (no trend)
  \caption{
  \textbf{Advantage of an adaptive success-bias weight.}
  Both success-bias weight $\alpha$ (\textbf{a}) and estimation error (\textbf{b}) decrease during the role-model choosing process, demonstrating that influence is becomes more advantageous as more copiers have made their choice.
However, when $\alpha$ is homogeneous, the mean estimation error doesn't decrease, regardless of $\alpha$ or $\eta$.
The mean estimation error in the homogeneous $\alpha$ model is larger by a factor of $10$ than the adaptive $\alpha$ model.
Here, population size $N=200$; estimation error is normally distributed $e \sim N(0,\eta^2)$ with standard deviation $\eta=$0.0001 (blue), 0.001 (orange), 0.01 (green), plots are average of $300$ simulations.}	
  \label{fig:influence_advantage}
\end{figure}

>>>>

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/final/GBD_validation.pdf}
  \caption{
  \textbf{Numerical validation for the GBD approximation.}
  The approximation (orange) fits simulation results (blue) well when using 1,000 simulations for both models.
  Here, population size, $N=2,000$; bias weight, $\alpha=0.1$; idea phenotype value, $\hat{A}=1$; role-model traits $\vec{A} \sim N(0,1)$; success bias value, $\beta(A)=0.956$.}	
  \label{fig:GBD}
\end{figure}

@@

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/final/DM_validation.pdf}
  \caption{
  \textbf{Goodness of fit of Dirichlet-Multinomial approximation to the model.}
  The difference between the DM approximation distribution to observed model distribution (a) is very small. The DM approximation (orange) and the simulation results (blue) are almost indistinguishable.
  The optimal negative log likelihood (b) is given when using the DM approximation (black vertical line), and is much better when compared to a shuffled vector of the DM distribution (blue bars).
  Here, population size, $N=100$; number of distributions averaged, $m=100$; phenotype values, $\hat{A}=1$, $A \sim N(0,1)$; homogeneous success bias weight, $\alpha=0.5$.}	
  \label{fig:DM_validation}
\end{figure}

@@

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/final/num_sims.pdf}
  \caption{
  \textbf{Minimum number of simulations required to get a good approximation.}
  The approximation (orange) fits simulation results (blue) well when using 1,000 simulations for both models. The approximated value (black) is given by equation \cref{eq:kimura_p}.
  Markers for average value across simulations, error bars for 95\% confidence interval.
  Here, population size, $N=1000$; bias weight, $\alpha=0.5$; phenotype values, $\hat{A}=1$, $A=0.7$; success bias value, $\beta(A)=0.956$.}	
  \label{fig:num_sims}
\end{figure}

\begin{figure}
    \includegraphics[width=\linewidth]{../figures/final/full_vs_dm_mutation.pdf}
  \caption{
  \textbf{Robustness of DM approximations to inclusion of estimation error.} Both the DM approximation (orange) and Kimura's equation (black line) fit the stochastic simulations (blue) well even with a high estimation error rate. Markers for average across simulations, error bars for 95\% confidence intervals.
  5,000 simulations per data point; population size, $N=1000$; bias weight, $\alpha=0.1$; phenotype values, $\hat{A}=1$,$A=0.7$; bias strength parameter $J\sim N(1,x^2)$ where $x \in [10^{-7},10^{-1}]$.
  }	
  \label{fig:hetro_error}
\end{figure}

\begin{figure}
    \includegraphics[width=\linewidth]{../figures/final/full_vs_dm_changing_alpha.pdf}
   \caption{\textbf{Robustness of DM approximations to variation in the bias weight $\alpha$.} 
   Both the DM approximation (orange) and Kimura's equation (black line) fit the stochastic simulations (blue) well even with a high variation in success bias weight.Markers for average across $5,000$ simulations, error bars are 95\% confidence intervals.
  Here, population size, $N=1000$; success bias weight normally distributed, $\alpha\sim N(0.5,x^2)$; phenotype values ,$\hat{A}=1$,$A=0.7$; success bias value, $\beta(A)=0.956$.}	
  \label{fig:hetro_alpha}
\end{figure}

@@

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/final/kimura_var.pdf}
  \caption{\textbf{Fixation probability and time approximation for varying bias weight and success coefficient.}
  Fixation probability (\textbf{A}) and time (\textbf{B}) for varying bias weight on the bottom x-axis. The corresponds effective population size is on the top x-axis.
  The approximation (black) fits DM simulations results (green) and Wright-Fisher simulation results (orange) when using \cref{eq:kimura_p}.
  Fixation probability (\textbf{A}) is limited at approximately $2s$ (blue), as in the classic WF model.
  Markers are averages of $10,000$ simulations, error bars show 95\% confidence intervals for (\textbf{A,B}) and 75\% for (\textbf{C,D}).
   Here, Population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=0.7$ (\textbf{A,B}), $A=a\cdot \hat{A}, a \in [0.01,0.99]$ (\textbf{C,D}); success coefficient, $1-\beta=s=0.044$ (\textbf{A,B}); bias weight, $\alpha=0.01$ (\textbf{C,D}).}
  \label{fig:var_alpha}
\end{figure}

@@

\begin{figure}[H]
    \includegraphics[width=\linewidth]{../figures/final/ch_env.pdf}
  \caption{\textbf{Sensitivity of fixation probability approximation in a changing environment to varying success coefficient and bias weight and robustness to varying ratios of cycles.}
\textbf{(A)} When the bias weight, or effective population size varies, the approximation (blue; \cref{eq:ch_env}) fits simulation results (orange). 
\textbf{(B)} When success bias is large ($1-\beta > 0.1$),  
simulation results (orange) are underestimated by the approximation (blue; \cref{eq:ch_env}). With even larger success bias ($1-\beta > 0.35$), even the constant environment approximation (green; \cref{eq:kimura_p}) underestimate simulation results, likely because the diffusion equation approximation assumes weak selection.
We can also see the robustness of the approximation to changes in cycle ratios (\textbf{C,D}): The approximation (blue) fits the simulations (orange) well for different sizes of the changing environment cycle, where $k$ is the number of generations where the bias is advantageous, and $l$ are the generations it is disadvantageous.
When $k<l$ the approximation is good. When $k>l$, the approximation and the simulations are both very close to the constant environment approximation (green). 
Markers show average of $10,000$ simulations, error bars show are 75\% \textbf{(A,C,D)} and 95\% \textbf{(B)} confidence intervals.
  Here, population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=0.9$ (\textbf{A,B}), $A=0.8$ \textbf{(C,D)}; In \textbf{(A)} success bias/selection coefficient is: $1-\beta=s=0.005$; In \textbf{(B,C,D)} success bias weight is: $\alpha=0.1$.}
  \label{fig:ch_env_alpha_beta}
\end{figure}

@@

\begin{figure}[H]
    \includegraphics[width=\linewidth]{../figures/final/choose_bias.pdf}
  \caption{
  \textbf{Advantage of an adaptive success-bias weight.}
  Both success-bias weight $\alpha$ (\textbf{a}) and estimation error (\textbf{b}) decrease during the role-model choosing process, demonstrating that influence is becomes more advantageous as more copiers have made their choice.
However, when $\alpha$ is homogeneous, the mean estimation error doesn't decrease, regardless of $\alpha$ or $\eta$.
The mean estimation error in the homogeneous $\alpha$ model is larger by a factor of $10$ than the adaptive $\alpha$ model.
Here, population size $N=200$; estimation error is normally distributed $e \sim N(0,\eta^2)$ with standard deviation $\eta=$0.0001 (blue), 0.001 (orange), 0.01 (green), plots are average of $300$ simulations.}	
  \label{fig:influence_advantage}
\end{figure}
